# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

First, let's get the data and check the dimensions of the data frame.

```{r GetData}
setwd("C:/Users/Owner/Documents/DataScienceCertification/RepoResearch/peer1/RepData_PeerAssessment1")

df1<-read.csv(unz("activity.zip", "activity.csv"), colClasses=c("numeric","Date","numeric"))

# Check Dim
dim(df1)


```

## What is mean total number of steps taken per day?

Now we plot a histogram of the total number of steps taken per day and   
calculate the associated mean and median... 

```{r AveStepsPerDay}

dfStepsPerDay<-aggregate(steps~date,data=df1,sum)
myMean<-summary(dfStepsPerDay$steps)[4]
myMedian<-summary(dfStepsPerDay$steps)[3]
summary(dfStepsPerDay$steps)
hist(dfStepsPerDay$steps,xlab="Steps Per Day",main="Histogram of Steps Per Day",col="blue")


```

The mean number of steps per day is `r myMean`.  
The median number of steps per day is `r myMedian`.


## What is the average daily activity pattern?

Here we simply calculate the mean number of steps for each time interval, generate  
a plot, and find the interval with the highest mean...

```{r MaxInterval}

dfStepsPerInterval<-aggregate(steps~interval,data=df1,mean)
maxStepsInterval<-dfStepsPerInterval[dfStepsPerInterval$steps==max(dfStepsPerInterval$steps),]
maxInterval<-maxStepsInterval[1]
maxSteps<-round(maxStepsInterval[2],0)


plot(dfStepsPerInterval$steps~dfStepsPerInterval$interval,type="l",col="blue",
     ylab="Mean Number of Steps",xlab="Interval",main="Mean Steps Per Interval")


```

The 5-minute interval, on average across all the days in the dataset, containing the maximum number of steps is interval `r maxInterval`.   
The mean number of steps in this interval is `r maxSteps`.


## Imputing missing values

Time to deal with missing values (NA).  First, how many are there?

```{r NumberMissing}

nMissing<-sum(is.na(df1))

```

The number of missing values in the dataset = `r nMissing`.  

  
Now impute missing values - keeping it simple by imputing means for each interval...  
Include a quick view of the data sets and dimensions to ensure successful imputation

```{r Impute}
dfMeans<-aggregate(steps~interval,data=df1,mean,rm.NA=TRUE)

df1NA<-df1[is.na(df1$steps),]
df2<-merge(df1NA[,c("date","interval")],dfMeans)
df3<-rbind(df1,df2)
df4<-df3[complete.cases(df3),]
df4<-df4[order(df4$date, df4$interval),]

head(df1)
head(df4)

dim(df1)
dim(df4)

```

After successfully imputing missing data, we plot a histogram of total steps per day

```{r HistImputed}
dfStepsPerDayImputed<-aggregate(steps~date,data=df4,sum)
hist(dfStepsPerDayImputed$steps,xlab="Steps Per Day",main="Histogram of Steps Per Day with Imputed Data",col="blue")



```

A quick check with the summary function suggests that imputing data changed the summary statistics for total steps per day very little.  Here is a summary on the original data...

```{r Summary}

summary(dfStepsPerDay$steps)


```


And here is a summary on the imputed data set...

```{r ImputeSummary}

summary(dfStepsPerDayImputed$steps)
myMeanImputed<-summary(dfStepsPerDayImputed$steps)[4]
myMedianImputed<-summary(dfStepsPerDayImputed$steps)[3]

```

Recall, the mean number of steps per day in the orignal data was `r myMean`.  
The mean number of steps per day in the imputed data is `r myMeanImputed`.  

Recall, the median number of steps per day in the orignal data was `r myMedian`.  
The median number of steps per day in the imputed data is `r myMedianImputed`.  

Therefore, imputing the data did not change the average total steps per day.  




## Are there differences in activity patterns between weekdays and weekends?

To do this, we need to identify and differentiate weekdays and weekends.  
First identify the day of the week, then split into weekdays and weekends.  
Check split with table function...

```{r weekend}

df4$Day.of.Week<-weekdays(df4$date)
df4$Split<-lapply(df4$Day.of.Week, function(x) ifelse(x %in% c("Saturday","Sunday"), "Weekend","Weekday"))

df4$Split<-factor(df4$Split,levels=c("Weekday","Weekend"))
table(df4$Split,df4$Day.of.Week)


```


We've got the data split by weekdays and weekends.  
Now make a plot of activity across both levels (weekday, weekend).   
To do so we must first get interval averages by level.  
Note we are using imputed dataset (df4)...

```{r intervalsByWeekend}

df5<-aggregate(steps~interval+Split,data=df4,mean)
dim(df5)

```

Looks like we have the data we need, so let's make the plot...

```{r panelPlot}
library(ggplot2)
pPanel<-ggplot(data=df5,aes(x=interval,y=steps))
pPanel<-pPanel + facet_wrap(~Split,nrow=2)
pPanel<-pPanel + geom_line(col="blue")
pPanel

```

A quick peek at the graphs suggests that this person gets up later on weekends  
(i.e. is less active in the early morning ~ 6-8 AM).  It also appears that he or she  
stays up a little later on weekends (i.e. is more active late in the day ~ 10 PM and after).  
Finally, this person is generally more active throughout the day on weekends.  He or she  
is probably a bright data scientist chained to the desk during the week, only getting up  
to go to meetings, the coffee pot, and then the restroom...   :)
